Model Comparison

### 2021-02-26 paper/supplementary_material.Rmd
clean repo for revision

---
title: "How young children integrate information sources to infer the meaning of words"
subtitle: "Supplementary material"
author: "Manuel Bohn, Michael Henry Tessler, Megan Merrick and Michael C. Frank"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
bibliography: library.bib
#csl: pnas.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\thefigure}{S\arabic{figure}}
---

@@

The goal of this study is to investigate the integration of information during word learning in children between 2 and 5 years of age. As a first step, we replicate earlier work showing that children rely on different forms of pragmatic and semantic information during word learning. In [Experiment 1](#experiment-1-mutual-exclusivity), we show that children make a so-called mutual exclusivity inference and that this inference depends on children's developing semantic knowledge. In [Experiment 2](#experiment-2-common-ground), we show that children make inferences about word meanings based on common ground. In [Experiment 3](#experiment-3-combination), we show that, when combined in one procedure, children are sensitive to the way that the two inferences are aligned.

@@

The first experiment tested the so-called mutual exclusivity inference in children between 2 and 5 years of age. The general inference can be described as follows: when presented with a familiar and an unfamiliar object, children expect a novel word to refer to the unfamiliar object [e.g. @markman1988children]. A range of explanations have been put forward for the cognitive basis of this inference [see @lewis2020-me for a discussion]. Here, we treat mutual exclusivity as a pragmatic phenomenon [e.g. @clark1987principle]. The inference process is specified in the model below.

@@

```{r fig1,echo = F,  fig.cap = "Schematic experimental procedure with screenshots from the experiments.", out.width="100%"}
knitr::include_graphics("../figures/method_overview.png")
```

@@

```{r}
#fig2 <- ggarrange(plot_aoa, plot_me, cor_plot_me,  labels = c("A","B","C"), nrow = 1, widths = c(1,1.3,1), common.legend = F)
#ggsave(plot = fig2, "../figures/figureS2.png", width = 12, height = 4)
```

@@

```{r fig2, include = T, fig.cap = "A:Familiar words and corresponding pictures by rated age of acquisition. B: Developmental trajectories of mututal exclusivity effect by familiar object based on the mean of the model posterior distribution. Dots show individual datapoints. Lighter colors indicate later rated age of acquisition. Dotted line indicates a level of performance expected by chance. C: Correlation between rated age of acquisiton and mutual exclusivity effect (model based intercept for each familiar object).", out.width="100%"}
knitr::include_graphics("../figures/figureS2.png")
```

@@

To estimate the parameter distributions, we collected samples from six independent MCMC chains, collecting 150,000 samples from each chain and removing the first 50,000 for burn-in. We excluded samples from one chain because it got stuck on a local maximum which resulted in parameter distributions that were substantially different from the other chains. To directly access the model predictions see the file `model_comparison.Rmd` in the online repository.

@@

```{r table5}
model_comp_pred <- readRDS("../saves/pred_model_comparison.rds")%>%
  mutate(model = factor(model),
         `Model comparison` = fct_recode(model, 
                            "integration vs no word knowledge" = "prag_vs_global",
                            "integration vs no common ground" = "prag_vs_flat",
                            "integration vs no speaker informativeness" = "prag_vs_prior",
                            "no word knowledge vs no common ground" = "global_vs_flat",
                            "no word knowledge vs no speaker informativeness" = "global_vs_prior",
                            "no common ground vs no speaker informativeness" = "flat_vs_prior"),
        `Bayes factor` = format(round(exp(logBF)), trim = T, justify = "right", scientific = F))%>%
  select(-logBF, -model)

@@

### Rational integration model

@@

### Biased model

@@

### Developmental bias model

@@

Perhaps unsurprisingly, we see that informing the parameters for semantic knowledge, speaker informativeness and common ground sensitivity by the data from Experiment 3 greatly improves the model fit (comparison: integration (explanation) vs integration (prediction)). We also see that the explanation **rational integration model** provides, by far, the best fit to the data compared to the two **biased models**, even though the rational integration model has fewer free parameters than the **biased model**. Interestingly, the prediction **rational integration model** also had a better fit, even though its parameters were not constrained by the data from Experiment 3. When comparing the two **biased models** directly, we see that an age sensitive bias parameter did not result in a substantially better fit, providing additional evidence that the integration process itself is not changing across development.

@@

```{r table6}
model_comp_pred <- readRDS("../saves/expl_model_comparison.rds")%>%
  mutate(model = factor(model),
         `Model comparison` = fct_recode(model, 
                            "integration (explanation) vs integration (prediction)" = "fb_vs_free",
                            "integration (explanation) vs biased" = "fb_vs_mixture",
                            "integration (prediction) vs biased" = "free_vs_mixture",
                            "integration (explanation) vs developmental bias" = "fb_vs_dev_mixture",
                            "integration (prediction) vs developmental bias" = "free_vs_dev_mixture",
                            "developmental bias vs biased" = "dev_mixture_vs_mixture"),
        `Bayes factor` = format(round(exp(logBF)), trim = T, justify = "right", scientific = F))%>%
  select(-logBF, -model)

@@

```{r figa1, include = T, cache=F,  fig.cap = "Posterior distribution of intercept term for semantic knowledge for each object by model.", fig.width=12, fig.height = 6}

item_comp_int <- item_comp %>%
  filter(parameter == "intercept")

ggplot(item_comp_int, aes(x = value, fill = model, col = model))+
  geom_density(alpha = 0.8, adjust = 2)+
  xlab("value")+
  facet_wrap( ~ item, nrow = 2, scales = 'free')+
  theme_few()+
  xlim(-10,5)+
  theme(legend.position = "bottom")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```

### Slope

```{r figa2, include = T, cache=F,  fig.cap = "Posterior distribution of slope term for semantic knowledge for each object by model.", fig.width=12, fig.height = 6}

item_comp_slope <- item_comp %>%
  filter(parameter == "slope")

ggplot(item_comp_slope, aes(x = value, fill = model, col = model))+
  geom_density(alpha = 0.8, adjust = 2)+
  xlab("value")+
  facet_wrap( ~ item, nrow = 2, scales = 'free_y')+
  xlim(-1,4)+
  theme_few()+
  theme(legend.position = "bottom")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```

@@

ggplot(global_comp, aes(x = value, fill = model, col = model))+
  geom_density(alpha = 0.8, adjust = 2)+
  xlab("value")+
  facet_grid(type ~ parameter, scales = 'free')+
  theme_few()+
  theme(legend.position = "right")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```

>>>>

---
title: "How young children integrate information sources to infer the meaning of words"
subtitle: "Supplementary material"
author: "Manuel Bohn, Michael Henry Tessler, Megan Merrick and Michael C. Frank"
output: 
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
bibliography: library.bib
#csl: pnas.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\thefigure}{S\arabic{figure}}
---

@@

The goal of this study is to investigate the integration of information during word learning in children between 2 and 5 years of age. As a first step, we replicate earlier work showing that children rely on different forms of pragmatic and semantic information during word learning. In [Experiment 1](#experiment-1-mutual-exclusivity), we show that children make a mutual exclusivity inference and that this inference depends on children's developing semantic knowledge. In [Experiment 2](#experiment-2-common-ground), we show that children make inferences about word meanings based on common ground. In [Experiment 3](#experiment-3-combination), we show that, when combined in one procedure, children are sensitive to the way that the two inferences are aligned.

@@

The first experiment tested the mutual exclusivity inference in children between 2 and 5 years of age. The general inference can be described as follows: when presented with a familiar and an unfamiliar object, children expect a novel word to refer to the unfamiliar object [e.g. @markman1988children]. A range of explanations have been put forward for the cognitive basis of this inference [see @lewis2020-me for a discussion]. Here, we treat mutual exclusivity as a pragmatic phenomenon [e.g. @clark1987principle]. The inference process is specified in the model below.

@@

```{r fig1,echo = F,  fig.cap = "Schematic experimental procedure with screenshots from the experiments.", out.width="100%"}
knitr::include_graphics("figures/method_overview.png")
```

@@

@@

```{r fig2, include = T, fig.cap = "A:Familiar words and corresponding pictures by rated age of acquisition. B: Developmental trajectories of mututal exclusivity effect by familiar object based on the mean of the model posterior distribution. Dots show individual datapoints. Lighter colors indicate later rated age of acquisition. Dotted line indicates a level of performance expected by chance. C: Correlation between rated age of acquisiton and mutual exclusivity effect (model based intercept for each familiar object).", out.width="100%"}
knitr::include_graphics("figures/figureS2.png")
```

@@

To estimate the parameter distributions and generate model predictions, we collected samples from six independent MCMC chains, collecting 150,000 samples from each chain and removing the first 50,000 for burn-in. We excluded samples from one chain because it got stuck on a local maximum which resulted in parameter distributions that were substantially different from the other chains. Model predictions can be downloaded via the following link: https://keeper.mpdl.mpg.de/d/7ae3805dba064bada8e5/. For the model comparison, we collected 150,000 samples (removing 50,000 for burn-in) from ten independent chains (file: `webppl/llh_model.wppl`, see also `stats/model_comparison.Rmd` in the online repository).

@@

```{r table5}
model_comp_pred <- readRDS("../saves/model_comparison.rds")%>%
  filter(!grepl("ias",model))%>%
  mutate(model = factor(model),
         `Model comparison` = fct_recode(model, 
                            "integration vs no word knowledge" = "prag_vs_global",
                            "integration vs no common ground" = "prag_vs_flat",
                            "integration vs no speaker informativeness" = "prag_vs_prior",
                            "no word knowledge vs no common ground" = "global_vs_flat",
                            "no word knowledge vs no speaker informativeness" = "global_vs_prior",
                            "no common ground vs no speaker informativeness" = "flat_vs_prior"),
        `Bayes factor` = format(round(exp(logBF)), trim = T, justify = "right", scientific = T))%>%
  select(-logBF, -model)

@@

@@

#### Biased model

@@

#### Developmental bias model

@@

When comparing models via their marginal likelihood, we again see that the **rational integration model** provides, by far, the best fit to the data compared to the two **biased models**, even though the rational integration model has fewer free parameters than the **biased models**. When comparing the two **biased models** directly, we see that an age sensitive bias parameter did not result in a substantially better fit, providing additional evidence that the integration process itself is not changing across development.

@@

```{r table6}
model_comp_exp <- readRDS("../saves/model_comparison.rds")%>%
  filter(grepl("ias",model))%>%
  mutate(model = factor(model),
         `Model comparison` = fct_recode(model, 
                            "integration vs biased" = "prag_vs_bias",
                            "integration vs developmental bias" = "prag_vs_devBias",
                            "developmental bias vs biased" = "devBias_vs_bias"),
        `Bayes factor` = format(round(exp(logBF)), trim = T, justify = "right", scientific = F))%>%
  select(-logBF, -model)

knitr::kable(model_comp_exp, caption = "Model comparison using Bayes factors computed based on the marginal likelihood of each model given the data.", digits = 2, align = "l")
```

@@

```{r figa1, include = T, cache=F,  fig.cap = "Posterior distribution of intercept and slope term for semantic knowledge for each object by model.", fig.width=12, fig.height = 5}

ggplot(item_comp, aes(x = value, fill = model, col = model))+
  geom_density(alpha = 0.3, adjust = 2)+
  xlab("value")+
  facet_grid( parameter~ item, scales = 'free')+
  theme_few()+
  theme(legend.position = "bottom")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```

@@

ggplot(global_comp, aes(x = value, fill = model, col = model))+
  geom_density(alpha = 0.3, adjust = 2)+
  xlab("value")+
  facet_grid(type ~ parameter, scales = 'free')+
  theme_few()+
  theme(legend.position = "right")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```

### 2021-02-26 stats/model_comparison.Rmd
clean repo for revision

```{r}
library(tidyverse)
library(readxl)
library(matrixStats)
library(data.table)

@@

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```
# Data

```{r}
data_comb_pre <- read_csv("../data/combination.csv")

aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% data_comb_pre$item )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

data_comb <- data_comb_pre%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa), 
         source = "Data")

d_model_comp <- data_comb%>%
  select(-age)%>%
  mutate(age = round(age_month,2),
         item = as.character(item))%>%
  select(alignment, item, age, correct)

```

# Model comparison prediction

@@

d_model_comp.summarized <- d_model_comp %>% 
  group_by(alignment, item, age) %>% 
  summarize(n=n(), k = sum(correct))

# join predictions with the data

model_comp <- bind_rows(

  readRDS("../saves/model_pred_prag.rds") %>%
    right_join(d_model_comp.summarized) %>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_global.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

   readRDS("../saves/model_pred_flat.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_prior.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)
) %>%
  mutate(
    log_like = dbinom(x = k, size = n, prob = pred, log = TRUE)
  ) %>%
  group_by(chain, model, iteration) %>%
  summarize(total_likelihood_sample = sum(log_like)) %>%
  group_by(chain, model) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)%>%
  mutate(prag_vs_global = pragmatic - global, # subtract likelihood for each predictions and model
         prag_vs_flat= pragmatic - flat,
         prag_vs_prior= pragmatic - prior,
         global_vs_flat = global - flat,
         global_vs_prior = global - prior,
         flat_vs_prior = flat - prior) %>%
  select(-pragmatic, -global, -flat, - prior) %>%
  gather(model, logBF, -chain)

```

# Model comparison explanation

d_model_comp.summarized <- d_model_comp %>% 
  group_by(alignment, item, age) %>% 
  summarize(n=n(), k = sum(correct))

model_like <- bind_rows(

  readRDS("../saves/model_pred_prag.rds") %>%
    right_join(d_model_comp.summarized) %>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_prag_fb.rds") %>%
    mutate(model = "fullBayes")%>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)

    ,readRDS("../saves/model_pred_mm.rds") %>%
    mutate(model = "mixture")%>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)

      ,readRDS("../saves/model_pred_dmm.rds") %>%
    mutate(model = "dev_mixture")%>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)
) %>%
  mutate(
    log_like = dbinom(x = k, size = n, prob = pred, log = TRUE)
  ) %>%
  group_by(chain, model, iteration) %>%
  summarize(total_likelihood_sample = sum(log_like))

# variation across chains

model_by_chain <- model_like %>%
  na.omit(chain)%>%
  group_by(model, chain) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)

saveRDS(model_by_chain, "../saves/expl_model_likelihood_chain.rds")

# compare models
model_comp <- model_like %>%
  na.omit(chain)%>%
  group_by(model) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)%>%
  mutate(fb_vs_free = fullBayes - pragmatic,
          fb_vs_mixture= fullBayes - mixture,
         fb_vs_dev_mixture= fullBayes - dev_mixture,
          free_vs_mixture = pragmatic - mixture,
         free_vs_dev_mixture = pragmatic - dev_mixture,
         dev_mixture_vs_mixture = dev_mixture - mixture
         ) %>%
  select(-pragmatic, -fullBayes, -mixture, -dev_mixture) %>%
  gather(model, logBF)

saveRDS(model_comp, "../saves/expl_model_comparison.rds")

```

>>>>

```{r}
library(tidyverse)
library(matrixStats)
library(ggthemes)

@@

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

# Model comparison

Model predictions can be downloaded via the following link: https://keeper.mpdl.mpg.de/d/7ae3805dba064bada8e5/
In the code below, model predictions are assumed to be in the sub-folder called `webppl/output/`.

@@

# sample <- bind_rows(
#   read_csv("../webppl/output/llh_model_chain1.csv")%>%mutate(chain = 1),
#   read_csv("../webppl/output/llh_model_chain2.csv")%>%mutate(chain = 2),
#   read_csv("../webppl/output/llh_model_chain3.csv")%>%mutate(chain = 3),
#   read_csv("../webppl/output/llh_model_chain7.csv")%>%mutate(chain = 4),
#   read_csv("../webppl/output/llh_model_chain8.csv")%>%mutate(chain = 5),
#   read_csv("../webppl/output/llh_model_chain9.csv")%>%mutate(chain = 6),
#   read_csv("../webppl/output/llh_model_chain10.csv")%>%mutate(chain = 7),
#   read_csv("../webppl/output/llh_model_chain11.csv")%>%mutate(chain = 8),
#   read_csv("../webppl/output/llh_model_chain12.csv")%>%mutate(chain = 9),
#   read_csv("../webppl/output/llh_model_chain13.csv")%>%mutate(chain = 10)
# )
# 
# sample_like <-  sample%>%
#   filter(a == "likelihood")%>%
#   rename(type = a,
#          model = b, 
#          alignment = c,
#          age = d,
#          item = e, 
#          log_like = f)%>%
#   mutate(age = as.numeric(age))%>%
#   group_by(chain, model, iteration) %>%
#   summarize(total_likelihood_sample = sum(log_like))
# 
# saveRDS(sample_like, "../saves/likelihood_sample.rds")

sample_like <- readRDS("../saves/likelihood_sample.rds")

# model_by_chain <- sample_like %>%
#   group_by(chain, model) %>%
#   summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
#   spread(model, marginal_likelihood)
# 
# saveRDS(model_by_chain, "../saves/model_likelihood_chain.rds")

model_by_chain <-  readRDS("../saves/model_likelihood_chain.rds")

# model_like <- sample_like %>%
#   group_by(model) %>%
#   summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample))
# 
# saveRDS(model_like, "../saves/model_likelihood.rds")

model_like <-  readRDS("../saves/model_likelihood.rds")

# model_comp <- model_like  %>%
#   group_by(model) %>%
#   spread(model, marginal_likelihood)%>%
#   mutate(prag_vs_global = pragmatic - global, # subtract likelihood for each predictions and model
#          prag_vs_flat= pragmatic - flat,
#          prag_vs_prior= pragmatic - prior_only,
#          prag_vs_bias = pragmatic - bias,
#          prag_vs_devBias = pragmatic - devBias,
#          devBias_vs_bias = devBias - bias,
#          global_vs_flat = global - flat,
#          global_vs_prior = global - prior_only,
#          flat_vs_prior = flat - prior_only) %>%
#   select(-pragmatic, -global, -flat, - prior_only, -bias, -devBias) %>%
#   gather(model, logBF)
# 
# saveRDS(model_comp, "../saves/model_comparison.rds")

model_comp <- readRDS( "../saves/model_comparison.rds")
```

# visualize log-likelihood distributions

```{r}
p_sample_like <- sample_like%>%
  mutate(model = fct_recode(model,
                            "integration" = "pragmatic",
                            "no word knowledge" = "global",
                            "no common ground" = "flat",
                            "no speaker informativeness" = "prior_only",
                            "biased" = "bias",
                            "developmental bias" = "devBias"),
         model = factor(model, levels = c("integration","no word knowledge","no common ground", "no speaker informativeness","biased", "developmental bias")))

p_model_like <- model_like %>%
  mutate(model = fct_recode(model,
                            "integration" = "pragmatic",
                            "no word knowledge" = "global",
                            "no common ground" = "flat",
                            "no speaker informativeness" = "prior_only",
                            "biased" = "bias",
                            "developmental bias" = "devBias"),
         model = factor(model, levels = c("integration","no word knowledge","no common ground", "no speaker informativeness","biased", "developmental bias")))

ggplot(p_sample_like, aes(x = total_likelihood_sample, col = model, fill = model))+
  geom_density(alpha = .5)+
  geom_vline(data = p_model_like, aes(xintercept = marginal_likelihood, col = model), lty = 2)+
  theme_minimal()+
  labs(x ="log likelihood (sample)", y = "")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```
